Smart Drill-Down: Interactive Summarization of Relational Tables

### 2014-11-21 TableSummarization/Paper/algorithms.tex
minor changes
changes to greedy sections

\section{Smart Drill-Down Algorithms}
\label{sec:algorithms}
We now describe online algorithms for implementing
the smart drill-down operator.

@@

\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
We reduce the well known NP-Hard problem called the {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set}, which will prove NP-Hardness for Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is given below: 
\begin{problem}\label{prob:maximum-coverage}
Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ... S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximises $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
\end{problem}
Thus the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible. We now reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}.

Then we construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns. Column $i$ of $T$ corresponds to set $S_i \in S$ and each tuple of $T$ corresponds to one $u \in U$. For each $u \in U$, the tuple $t_u$ corresponding to $u$ defined as follows: For each $S_i \in S$, $$t_u(i) = 1 \textrm{ if } u \in S_i \textrm{ else } t_u(i) = 0$$

@@

Then a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance. This completes our reduction.

\subsection{Algorithm Overview}\label{sec:alg-overview}
We now present our algorithms for approximating the solution to Problem~\ref{prob:optimal-rule-set}. The problem consists of finding a set of given size $k$, that maximises Score. We first show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm. We start with an empty rule set $R$, and then run the greedy algorithm for $k$ steps, where in each step, we add the rule $r$ to $R$ that maximises Score($R \cup \left\lbrace r \right\rbrace$). In order to find the rule $r$ to add in each step, we need to measure $Score(R \cup \left\lbrace r \right\rbrace) - Score(R)$ for each $r$. This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent itemset mining.

In addition, the greedy algorithm can be run over a sample of the table instead of over the entire table, to reduce computation time for large tables. We later describe a scheme for maintaining multiple samples in memory and using them to improve response time for different drill down operations performed by the user.

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

@@

Now consider the greedy algorithm:

\begin{framed}
\vspace{-15pt}
\begin{enumerate}
\item Set $S = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
\item $S = S \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\vspace{-15pt}
\end{framed}

Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}. The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.

@@

In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one. A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.

@@

\subsection{Sampling for big tables}
\subsubsection{Sampling Overview}
We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory. First in Section~\ref{sec:sample-using}, we motivate the need for sampling. Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk. Finally, we mention some additional optimizations we can make, and heuristics for setting the minimum sample size required from the SampleHandler.

>>>>

\section{Smart Drill-Down Algorithms}
\label{sec:algorithms}
We now describe online algorithms for implementing
the smart drill-down operator. We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand). We further discuss bucketization of numerical attributes in Section~\ref{sec:extensions}.

@@

\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
We reduce the well known NP-Hard {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set};
thus demonstrating the NP-Hardness of Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is as follows: 
\begin{problem}\label{prob:maximum-coverage}
Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ... S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximises $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
\end{problem}
Thus, the goal of the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible. We now reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}.

\begin{proof}
We construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns. Column $i$ of $T$ corresponds to set $S_i \in S$ and each tuple of $T$ corresponds to one $u \in U$. For each $u \in U$, the tuple $t_u$ corresponding to $u$ defined as follows: For each $S_i \in S$, $$t_u(i) = 1 \textrm{ if } u \in S_i \textrm{ else } t_u(i) = 0$$

@@

Then, a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance. This completes our reduction.
\end{proof}

\subsection{Algorithm Overview}\label{sec:alg-overview}
Given that the problem is NP-Hard, we now present our algorithms 
for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
The problem consists of finding a set of given size $k$, that maximizes 
Score.

The next few sections fully develop the
details of our solution:

\squishlist
\item We show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm. At a high level, this greedy algorithm is simple to state. The algorithm runs for $k$ steps;
we start with an empty rule set $R$, and then at each step, we add the next best rule that maximizes Score

\item In order to find the rule $r$ to add in each step, we need to measure the impact on Score for each $r$. This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent item-set mining. 
\squishend

In some cases, the dataset may still be too large for us to return a good rule set in
a reasonable time; in such cases, we may want to run our algorithm on a sample of the table
rather than the entire table. In Section~\ref{sec:sampling}, we describe a scheme 
for maintaining multiple samples in memory and using them to improve response 
time for different drill down operations performed by the user.

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
\stitle{Submodularity:} We will now show that the Score function over sets of rules has a property called {\em submodularity}, giving us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

@@

\stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
as listed below, has desirable approximation guarantees:
\begin{framed}
\vspace{-15pt}
\begin{enumerate}
\item Set $S = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
\item $S = S \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\vspace{-15pt}
\end{framed}
Since Score is a submodular function of the set $S$, 
this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

The expensive step in the above procedure is the step where the Score is computed for every
single rule. Given the number of rules can be very large, this can be an especially time-consuming process.

Instead of using the procedure described above directly, we instead develop a ``parameterized'' version 
that will admit further approximation (depending on the parameter) in order to reduce computation further. We describe
this algorithm next.

\stitle{Parametrized Algorithm:} Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}. The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.

@@

\stitle{Finding the Best Marginal Rule:} In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one. A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.

@@

\subsection{Sampling for big tables}\label{sec:sampling}
\subsubsection{Sampling Overview}
We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory. First in Section~\ref{sec:sample-using}, we motivate the need for sampling. Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk. Finally, we mention some additional optimizations we can make, and heuristics for setting the minimum sample size required from the SampleHandler.
